---
title: "Aides Drug"
author: "Junior Cruz"
date: "Summer 2024"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

We are going to get a subset of data that gives us the drug users. 

```{r}
DrugUser<-filter(triUmp,DrugUser=="yes")
```

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")

```


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->



## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")

```

For the very first chart showing not many people got aids which is a typical situation because in a clinical trial not many people get what the supposed treatment prevent. In the second chart the amount of people in the tan are of this chart took the placebo while the blue took drugs, those who got aids 3x as many took the placebo than the prevented drugs. 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

Those who ended up infected 24% had taken the drug while 75% had taken the placebo, 3x more likely to get infected had taken the placebo than had taken the drug which is supposed to prevent them from getting infected, therefore these help prevent aids.  

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The fisher test for the count data shows that, the p value is .0000000423 I will reject the null hypothesis mainly because their is a big difference in taking the drug or placebo. This shows us that the people taking the drugs are being protected from aids we just do not know how the drug works. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

We conducted a series of test to see if the drugs or placebo to be effective of preventing aids. For the first chart showing that not many people got aids and in a clinical trial showing that not many people got what the supposed treatment was to prevent. In the second chart that we explored the amount of people in the tan area took the placebo while blue got the drugs, those who had aids three times as many took the placebo than the prevented drugs. Those who ended up infected 24% had taken the drug while 75% had taken the placebo, the three times more likely to get infected had taken the placebo than taking the drug which prevents them from getting infected and helps prevent aids. Now for the results of the data has shown that P value is shown to be .0000000423 which we have rejected the null hypothesis because of the huge difference in taking either the drug or the placebo. Therefore showing that those taking the drug seem to prevent aids although we have no idea how this drug works.